---
title: 'COPE data for NYU Langone cohort: New mom questionnaire'
author: "Michelle.VanTieghem"
date: "June 2, 2020"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  word_document:
    toc: yes
---


# README 
Includes all participants who completed New Mom questionnaire at either baseline or Week 2 follow-up. \ Questions were optional so not all Qs have same  # of participants. \
Data collected between 04-04-2020 and 06-02-2020. \

```{r, echo = F, message = F, warning = F, include = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F, include= F)

library(tidyverse)
library(ggplot2)
library(readxl)
library(kableExtra)
```

# load data
```{r, echo = F, include = F}
load("clean_data/new_mom_df_clean.Rdata")
N_completed_maternal_Q <- nrow(mom_data_all)

```

Total new moms completing questionnaire: `r N_completed_maternal_Q`

# Maternal ratings of general perinatal health care
## Current support 
How well have you been supported by your pre- and postnatal care provider(s) ?
```{r, echo = F, include = F}
support_table <- mom_data_all %>%
  filter(!is.na(mom_care_support_change)) %>%
  summarize(N = n(), 
           "Very well supported" = sum(mom_care_support == 1, na.rm = T),
           "Somewhat well supported" = sum(mom_care_support == 2, na.rm = T), 
           "Not very well supported"= sum(mom_care_support == 3, na.rm = T))

support_table_long <- support_table %>%
  gather(key = "Support_level", value = "N_endorsed",  -N) %>%
  mutate(Percent = round((N_endorsed/N) *100, 2)) %>%
  select(Support_level, N_endorsed, Percent)
```

```{r, echo = F}
(support_table_long)

```


## Change in perinatal care due to covid-19 outbreak
Did the support you received from your pre- and postnatal care provider(s) change due to the COVID-19 outbreak?
```{r, echo = F, include = F}

support_table <- mom_data_all %>%
  filter(!is.na(mom_care_support_change)) %>%
  summarize(N = n(), 
           "Significantly worsened" = sum(mom_care_support_change == 1, na.rm = T),
           "Somewhat worsened" = sum(mom_care_support_change == 2, na.rm = T), 
           "No change"= sum(mom_care_support_change == 3, na.rm = T), 
           "Not relevant" = sum(mom_care_support_change == 4, na.rm = T), 
           "Somewhat improved" = sum(mom_care_support_change == 5, na.rm = T), 
           "Significantly improved" = sum(mom_care_support_change == 6, na.rm = T))


support_table_long <- support_table %>%
  gather(key = "change", value = "N_endorsed",  -N) %>%
  mutate(Percent = round((N_endorsed/N) *100, 2)) %>%
  select(change, N_endorsed, Percent)
```

```{r, echo = F}
(support_table_long)

```


```{r, echo = F}

support_table_long %>%
  mutate(change = reorder(change, Percent)) %>%
  filter(change != "Not relevant") %>%
ggplot(aes(x = change, y = Percent)) + 
  geom_bar(stat =  "identity") + 
  ylim(0, 100) + coord_flip()  + theme_bw() + xlab("") +
  scale_fill_manual(values=c("dark red", "dark gray")) +
   theme(legend.position = c(0.8, 0.2)) + ylab ("Percent") +
  labs( title = "Change in perinatal care support due to Covid-19", 
 subtitle = "Percent of women reporting changes in pre or postnatal care", 
          caption = "Data source: NYU Langone COPE Study") 

```



# Changes to Birth plan - before and after covid
 0 = Delivery date before March 15, 2020 (before covid) \
 1 = Delivery date on or after March 15, 2020 (after covid) \
(question is check all that apply)
```{r, echo =F, include = F}

# get N for how many people answered this Q, to calculate %
N_total <- sum(!is.na(mom_data_all$any_birth_change))

# make table of changes
birth_changes_table <- mom_data_all %>%
  # only include people who answered this Q
  filter(!is.na(any_birth_change)) %>%
  # group by birth before or after March 15, 2020
  group_by(Delivery_date) %>%
  summarize(N_total = n(), 
            'Any change to birth plan' = sum(any_birth_change, na.rm = T),
            'Reduced access to preferred medications' = sum(birth_plan_changes___1 == 1, na.rm = T),
            'Changes to birth location' = sum(birth_plan_changes___2 == 1, na.rm = T),
            'Changes to birth schedule' = sum(birth_plan_changes___3 == 1 | 
                                                birth_plan_changes___4 == 1, na.rm = T),
            "Health care provider was not available" = sum(birth_plan_changes___5 == 1, na.rm = T),
            "Support person not permitted at delivery"= sum(birth_plan_changes___6 == 1, na.rm = T),
            'Separated from baby after delivery' = sum(birth_plan_changes___7 == 1 |
                                                         birth_plan_changes___8 == 1, na.rm = T),
            'Other change to birth plan' = sum(birth_plan_changes___10 == 1, na.rm = T),
            'No changes' = sum(birth_plan_changes___9 == 1, na.rm = T))

birth_changes_table_long <- birth_changes_table %>%
  gather(key = "Birth_Changes", value = "N_endorsed", -Delivery_date, -N_total) %>%
  select(Birth_Changes,  Delivery_date, N_endorsed, N_total) %>%
  mutate(Percent = round((N_endorsed/N_total)*100, 2)) %>%
  select(Birth_Changes, Delivery_date, N_endorsed, Percent)
save(birth_changes_table_long, file = "tables/birth_changes_table_long.Rdata")
```


## Change vs. no change to birth plan, by delivery date
```{r, echo = F}
(birth_changes_table_long %>%
   filter(Birth_Changes == "Any change to birth plan" |
            Birth_Changes == "No changes"))
```

## Types of changes to birth plan due to covid-19
```{r, echo = F}
(birth_changes_table_long %>%
   filter(Birth_Changes != "Any change to birth plan" &
            Birth_Changes != "No changes" & Delivery_date == "After March 15, 2020")) %>%
  select(Birth_Changes, N_endorsed, Percent)
```


```{r,  echo = F}
birth_plot <- birth_changes_table_long %>%
      filter(Delivery_date == "After March 15, 2020" & 
               Birth_Changes != "No changes") %>%
      mutate(Birth_Changes = reorder(Birth_Changes, Percent)) %>%
      ggplot(aes(x = Birth_Changes, y = Percent)) +
      geom_bar(stat =  "identity", position = position_dodge(), fill = "dark red") + 
      coord_flip()  + theme_bw()  + ylim(0, 100)+
           ylab("Percent of women who reported changes") +
      labs( title = "Birth & Delivery during the COVID-19 outbreak", 
            caption = "Data aggregated from 147 women who delivered after March 15th, 2020")  + 
        xlab ("") 
    
    # Display the plot (whether it has facets or not)
    birth_plot
```

  
# Changes in Postnatal health care due to covid-19 outbreak
(check all that apply)
```{r, echo = F, include = F}
# get N for how many people answered this Q, to calculate %
N_total <- sum(!is.na(mom_data_all$any_postnatal_change))

# make table of changes
postnatal_changes_table <- mom_data_all %>%
  # only include people who answered this Q
  filter(!is.na(any_postnatal_change)) %>%
  summarize(N_total = n(), 
            'Any change to postnatal care plan' = sum(any_postnatal_change, na.rm = T),
             'Family/friends unable to visit' = sum(postnatal_plan_changes___1 == 1, na.rm = T),
            'No access to lactation support' = sum(postnatal_plan_changes___2 == 1, na.rm = T),
            'Cancelled postpartum visit' = sum(postnatal_plan_changes___3 == 1, na.rm = T),
            'Virtual postpartum visit' = sum(postnatal_plan_changes___4 == 1, na.rm = T),
            "Unable to access preferred contraception" = sum(postnatal_plan_changes___5 == 1, na.rm = T),
            "Unable to talk about baby blues"= sum(postnatal_plan_changes___6 == 1, na.rm = T),
            'Virtual Baby well visits' = sum(postnatal_plan_changes___7 == 1, na.rm = T),
            'Cancelled Baby well visits' = sum(postnatal_plan_changes___8 == 1, na.rm = T),
            'Postponed immunizations for baby' = sum(postnatal_plan_changes___9 == 1, na.rm = T), 
            "Other change" =sum(postnatal_plan_changes___11 == 1, na.rm = T),
            "No change" = sum(postnatal_plan_changes___10 == 1, na.rm = T))

postnatal_changes_table_long <- postnatal_changes_table %>%
  gather(key = "Postnatal_Changes", value = "N_endorsed", -N_total) %>%
  select(Postnatal_Changes,N_endorsed, N_total) %>%
  mutate(Percent = round((N_endorsed/N_total)*100, 2)) %>%
  select(Postnatal_Changes,  N_endorsed, Percent)
```


## Change vs. no change to postnatal care plan, by delivery date
```{r, echo = F}

postnatal_changes_table_long %>%
   filter(Postnatal_Changes == "Any change to postnatal care plan" |
            Postnatal_Changes == "No change")
```

## Types of changes to birth plan due to covid-19
```{r, echo = F}
postnatal_changes_table_long %>%
   filter(Postnatal_Changes != "Any change to postnatal care plan" &
            Postnatal_Changes != "No change" ) %>%
  select(Postnatal_Changes, N_endorsed, Percent)

save(postnatal_changes_table_long, file = "tables/postnatal_changes_table_long.Rdata")
```

```{r}
# create the plot
 postnatal_plot <- postnatal_changes_table_long %>%
      filter(Postnatal_Changes != "No change") %>%
      mutate(Postnatal_Changes = reorder(Postnatal_Changes, Percent)) %>%
      ggplot(aes(x = Postnatal_Changes, y = Percent)) +
      geom_bar(stat =  "identity", position = position_dodge(), fill = "dark red") + 
      coord_flip()  + theme_bw() + ylab("Percent of women who reported changes") + 
       ylim(0, 100) + xlab("") +
      labs( title = "Postnatal care during the COVID-19 outbreak", 
            caption = "Data aggregated from 406 women with infants under 6 months of age")
postnatal_plot
```
