---
title: "covid testing and symptoms over time"
author: "Michelle.VanTieghem"
date: "July 14, 2020"
output:
  html_document:
    number_sections: no
    df_print: kable
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

# README

```{r, echo = F, message = F, warning = F, include = F}
knitr::opts_chunk$set(message = F, warning = F)

library(psych)
# load packages
library(corrplot)
library(tidyverse)
library(ggplot2)
library(readxl)

# load custom packages
source("../custom_functions.R")

## special colors for testing positive /neg
my_test_fill = scale_fill_manual(values = c("dark red", "dark blue", "blue", "red"))
my_test_color = scale_color_manual(values = c("dark red", "dark blue"))

```


# load data
```{r}
covid_long <- read.csv( "../clean_data/longitudinal_covid_test_symp.csv")

# create date bins
covid_long <- covid_long %>%
  mutate(date_of_visit = as.character(date_of_visit),
        date_bin = as.factor(ifelse(date_of_visit < "2020-05-01", "04", 
                           ifelse(date_of_visit < "2020-06-01", "05", 
                                  ifelse(date_of_visit < "2020-07-01", "06", 
                                         ifelse(date_of_visit < "2020-08-01", "07", NA)))))) %>%
  filter(!is.na(date_bin)) %>%
  mutate(timepoint = as.factor(ifelse(redcap_event_name == "covid_arm_1", 1, 
                            ifelse(redcap_event_name == "covid_week_2_follo_arm_1", 2, 
                                   ifelse(redcap_event_name == "covid_week_4_follo_arm_1", 3, 
                                   ifelse(redcap_event_name == "covid_week_6_follo_arm_1", 4, NA))))),
         group = as.factor(ifelse(status == "New Mom" & first_child == "N", "New mom, with older children", 
                        ifelse(status == "New Mom" & first_child == "Y", "New mom, first child",
                               ifelse(status == "Pregnant" & first_child == "Y", "Pregnant, first child", ifelse(status == "Pregnant" & first_child == "N", "Pregnant, with older children", NA))))))

```


# Calculating Testing Rates 
## across all visits - any test? 
 if more than one, use positive test
```{r}

covid_self_test_any <- covid_long %>%
  select(record_id, status, first_child, timepoint, self_test) %>%
  mutate(any_test = ifelse(is.na(self_test) | 
                             self_test == 1 |
                             self_test == 999, 0, 1)) %>%
  spread(key = timepoint, value = any_test) %>%
  rename(Test1 = '1', Test2 = '2', Test3 = '3', Test4 ='4') %>%
  mutate(Num_tests = rowSums(cbind(Test1, Test2, Test3, Test4), na.rm = T), 
         any_test = ifelse(Num_tests > 0, 1, 0), 
         no_test = ifelse(Num_tests == 0, 1, 0)) %>%
  select(-self_test)

## NOT THE RIGHT LENGTH!!
nrow(covid_self_test_any)
length(unique(covid_self_test_any$record_id))

# get test results... 
covid_self_test_result <- covid_long %>%
  select(record_id, timepoint, self_test) %>%
  mutate(any_test = ifelse(is.na(self_test) | 
                             self_test == 1 |
                             self_test == 999, 0, 1),
         positive_test = ifelse(any_test == 0, NA,
                                ifelse(self_test == 2, 1, 
                                ifelse(self_test == 3, 0, NA)))) %>%
  spread(key = timepoint, value = positive_test) %>%
  rename(Pos1 = '1', Pos2 = '2', Pos3 = '3', Pos4 ='4') %>%
  # add up all positive tests per subject
  mutate(Num_pos_tests = rowSums(cbind(Pos1, Pos2, Pos3, Pos4), na.rm =T),
         never_tested = ifelse(is.na(Pos1) & is.na(Pos2) & is.na(Pos3) & is.na(Pos4), 1, 0),
         # if they have never been tested, make this NA
         # if they have been tested, mark if number posiitve is more than zero.
         any_pos_test = ifelse(never_tested == 1, NA, 
                               ifelse(Num_pos_tests > 0, 1, 0))) %>%
  select(-self_test)
nrow(covid_self_test_result)


# combine tables 
wide_test_data <- covid_self_test_any%>%
  merge(.,  covid_self_test_result , by = "record_id")


self_test_any_table <- covid_self_test_any %>%
  group_by(status) %>%
  summarize(n = n(), 
            any_test = sum(any_test), 
            any_test_percent = round((any_test/n)*100, 2))

self_test_result_table <- covid_self_test_result %>%
  group_by(status) %>%
  summarize(n = n(), 
            any_pos_test = sum(any_pos_test, na.rm = T), 
            any_pos_test_percent = round((any_pos_test/n)*100,2))
  
## add to table with any test rate
cbind(self_test_result_table, self_test_any_table$any_test, self_test_any_table$any_test_percent)

```


## plot rates of test restults. 
```{r}
self_test_table %>%
  ggplot(aes(x = status, y = percent)) +
  geom_col(aes(fill = status)) + 
  my_colors_fill + 
  custom_theme + ylim(0, 100) +
  labs(title = "Percent of women reporting Covid-19 testing", 
       y = "Percent", x = "", 
       caption = paste("Data aggregated from", sum(self_test_table$n), "new and expecting mothers"))
  

```


# Testing over time...
## recode testing and symptom variables 
```{r}
test_vars <- covid_long %>% 
  select(record_id, timepoint, contains("test")) %>%
  na_if(., '999') %>% # change 999 to NAs
  # for each variable, recode the numbers to factors:
  mutate_if(is.numeric, funs(recode(., '1' = 'Never tested', '2' = 'Positive', 
                                    '3' = 'Negative', '4' = 'waiting')))
symp_vars <- covid_long %>% 
  select(record_id, timepoint, contains("symp")) %>%
  na_if(., '999') %>% # change 999s to NAs
   mutate_if(is.numeric, funs(recode(., '1' = 'No symptoms', '2' = 'Current symptoms', 
                                    '3' = 'Past symptoms')))

covid_long <- covid_long %>%
  # remove the original versions
  select(-contains("test"), -contains("symp")) %>%
  # add the new versions
  left_join(., test_vars, by = c("record_id", "timepoint")) %>%
  left_join(., symp_vars, by = c("record_id", "timepoint"))
  
```


## Testing rate 
percent tested 

## Percent positive 
of entire sample, or of tested only? 
are the same people reporting their positive test over and over? or represents new tests over time?
same thing for family... 


## how tests per time point? 
```{r}
N_per_tp <- covid_long %>%
  group_by(timepoint) %>% 
  count()

covid_test_table <- covid_long %>% 
  group_by(timepoint, self_test) %>%
  count() %>%
  mutate(Percent = round((n/ N_per_tp$n[N_per_tp$timepoint == timepoint])*100, 2)) %>%
  ungroup()

covid_test_table

N_tested <- 
```

## plot testing data 
```{r}

covid_test_table %>%
  filter(self_test != "waiting" & !is.na(self_test) & self_test != "Never tested") %>%
  mutate(self_test = reorder(self_test, Percent)) %>%
  ggplot(aes(x = timepoint, y = Percent, fill = self_test)) + 
  geom_bar(stat = "identity") + my_colors_fill +
  custom_theme + xlab ("Survey Time Point") + ylim(0, 50) +
  labs(fill = 'Covid-19 Test Result', 
        title = "Self-reported Covid-19 Test Results over time", 
        caption = paste("Data aggregated acrossnew and expecting mothers"))

```


## plot testing data 
```{r}

covid_test_table %>%
  filter(self_test != "waiting" & !is.na(self_test) & self_test != "Never tested") %>%
  mutate(self_test = reorder(self_test, Percent)) %>%
  ggplot(aes(x = timepoint, y = Percent, fill = self_test)) + 
  geom_point(aes(color = self_test)) +
  geom_line(aes(group = self_test, color = self_test)) + 
  theme_bw() + my_colors_color

```
