---
title: "COPE EDA: emtoions & feelings over time"
author: "Michelle.VanTieghem"
date: "July 14, 2020"
output:
  html_document:
    number_sections: no
    df_print: kable
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

# README
bsi scores have been transformed to 0-6 so that 0 = no endorsement (instead of 1-7)
only including variables that were repeated across all events, 
so not the entire BSI that was collected at baseline.
Note: all of this code is visualization of data, but no statistical models = no inference! 

```{r, echo = F, message = F, warning = F, include = F}
knitr::opts_chunk$set(message = F, warning = F)

library(psych)
# load packages
library(corrplot)
library(tidyverse)
library(ggplot2)
library(readxl)

# load custom packages
source("../custom_functions.R")
```


# load data
```{r}
emo_long <- read.csv( "../clean_data/longitudinal_bsi_ptsd.csv")

# create date bins
emo_long <- emo_long %>%
  mutate(date_of_visit = as.character(date_of_visit)) %>%
  mutate(date_bin = ifelse(date_of_visit < "2020-05-01", "04", 
                           ifelse(date_of_visit < "2020-06-01", "05", 
                                  ifelse(date_of_visit < "2020-07-01", "06", 
                                         ifelse(date_of_visit < "2020-08-01", "07", NA))))) %>%
  filter(!is.na(date_bin)) %>%
  mutate(timepoint = ifelse(redcap_event_name == "covid_arm_1", 1, 
                            ifelse(redcap_event_name == "covid_week_2_follo_arm_1", 2, 
                                   ifelse(redcap_event_name == "covid_week_4_follo_arm_1", 3, 
                                   ifelse(redcap_event_name == "covid_week_6_follo_arm_1", 4, NA)))),
         group = ifelse(status == "New Mom" & first_child == "N", "New mom, with older children", 
                        ifelse(status == "New Mom" & first_child == "Y", "New mom, first child",
                               ifelse(status == "Pregnant" & first_child == "Y", "Pregnant, first child", ifelse(status == "Pregnant" & first_child == "N", "Pregnant, with older children", NA)))))

```

## compute bsi scores and PTSD scores 
```{r}
bsi_vars <- emo_long %>% 
  select(starts_with("bsi"))

ptsd_vars <- emo_long %>% 
  select(starts_with("ptsd"))

emo_long <- emo_long %>%
  mutate(bsi_ave = rowMeans(bsi_vars, na.rm =T), 
         ptsd_ave = rowMeans(ptsd_vars, na.rm = T))
```

## how many per group?
```{r}

emo_long %>% 
  group_by(first_child, status) %>%
  count()

```

## plot raw data 
with a smoothing function - NOTE, just FOR VISUALIZATION - not modeled!! 
```{r}
emo_long %>%
  ggplot(aes(x = timepoint, y = bsi_ave)) + 
 geom_line(aes(color = status, group = record_id), alpha = 0.2) +
 theme_bw() + xlab ("survey timepoint") + ylab ("Mean BSI score") +
  geom_smooth(aes(group = status, color = status, fill = status))

```


## calculate scores by survey time points 
```{r}

bsi_table <- emo_long %>%
  group_by(timepoint, status) %>%
  summarize(bsi_mean = mean(bsi_ave, na.rm = T), 
         bsi_sd = sd(bsi_ave, na.rm = T), 
         bsi_upper = bsi_mean + bsi_sd, 
         bsi_lower = bsi_mean - bsi_sd, 
         ptsd_mean = mean(ptsd_ave, na.rm = T), 
         ptsd_sd = sd(ptsd_ave, na.rm = T),
         ptsd_upper = ptsd_mean + ptsd_sd, 
         ptsd_lower = ptsd_mean - ptsd_sd) 
  

bsi_table
```


## BSI plot over time, by survey time points 
error bars = 1 SD 
```{r}
bsi_table %>%
  ggplot(aes(x = timepoint, y = bsi_mean)) + 
  geom_ribbon(aes(ymin = bsi_lower, ymax = bsi_upper, fill = status), alpha = 0.1) +
  geom_line(aes(color = status, group = status)) +
 theme_bw() + xlab ("survey timepoint") + ylab ("Mean BSI score") 

```

## calculate scores by actual date bins
```{r}
# create date bins 

bsi_table2 <- emo_long %>%
  group_by(date_bin,  status) %>%
  summarize(bsi_mean = mean(bsi_ave, na.rm = T), 
         bsi_sd = sd(bsi_ave, na.rm = T), 
         bsi_upper = bsi_mean + bsi_sd, 
         bsi_lower = bsi_mean - bsi_sd,
              ptsd_mean = mean(ptsd_ave, na.rm = T), 
         ptsd_sd = sd(ptsd_ave, na.rm = T),
         ptsd_upper = ptsd_mean + ptsd_sd, 
         ptsd_lower = ptsd_mean - ptsd_sd)

bsi_table2
```


## BSI plot over time, by actual date 
error bars = 1 SD 
```{r}
bsi_table2 %>%
  ggplot(aes(x = date_bin, y = bsi_mean)) + 
  geom_line(aes(color = status, group = status)) +
 theme_bw() + xlab ("Month") + ylab ("Mean BSI score") +
  geom_errorbar(aes(ymin = bsi_mean - bsi_sd, ymax = bsi_mean + bsi_sd), alpha = 0.1) 

```