---
title: 'COPE data for NYU Langone cohort: Delivery / Birth changes'
author: "Michelle.VanTieghem"
date: "June 2, 2020"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  word_document:
    toc: yes
---


# README 
Includes all participants who completed New Mom questionnaire at either baseline or Week 2 follow-up. \ Questions were optional so not all Qs have same  # of participants. \
Data collected between 04-04-2020 and 06-02-2020. \

```{r, echo = F, message = F, warning = F, include = F}
knitr::opts_chunk$set(message = F, warning = F)

# load packages
library(tidyverse)
library(ggplot2)
library(readxl)
library(lubridate)

# load custom packages
source("custom_functions.R")


load("clean_data/new_mom_df_clean.Rdata")
(N_completed_maternal_Q <- nrow(mom_data_all))

```


# Changes to Birth plan - before and after covid
 0 = Delivery date before March 15, 2020 (before covid) \
 1 = Delivery date on or after March 15, 2020 (after covid) \
(question is check all that apply)
```{r, echo =F, include = F}

# get N for how many people answered this Q, to calculate %
N_total <- sum(!is.na(mom_data_all$any_birth_change))

# make table of changes
birth_changes_table <- mom_data_all %>%
  filter(covid19_survey_for_new_mothers_complete ==2) %>%
  # only include people who answered this Q 
  filter(!is.na(any_birth_change)) %>%
  # group by birth before or after March 15, 2020
  group_by(Delivery_date) %>%
  summarize(N_total = n(), 
            'Any change to birth plan' = sum(any_birth_change, na.rm = T),
            'Reduced access to preferred medications' = sum(birth_plan_changes___1 == 1, na.rm = T),
            'Changes to birth location' = sum(birth_plan_changes___2 == 1, na.rm = T),
            'Changes to birth schedule' = sum(birth_plan_changes___3 == 1 | 
                                                birth_plan_changes___4 == 1, na.rm = T),
            "Health care provider was not available" = sum(birth_plan_changes___5 == 1, na.rm = T),
            "Support person not permitted at delivery"= sum(birth_plan_changes___6 == 1, na.rm = T),
            'Separated from baby after delivery' = sum(birth_plan_changes___7 == 1 |
                                                         birth_plan_changes___8 == 1, na.rm = T),
            'Other change to birth plan' = sum(birth_plan_changes___10 == 1, na.rm = T),
            'No changes' = sum(birth_plan_changes___9 == 1, na.rm = T))

birth_changes_table_long <- birth_changes_table %>%
  gather(key = "Birth_Changes", value = "N_endorsed", -Delivery_date, -N_total) %>%
  select(Birth_Changes,  Delivery_date, N_endorsed, N_total) %>%
  mutate(Percent = round((N_endorsed/N_total)*100, 2)) %>%
  select(Birth_Changes, Delivery_date, N_endorsed, Percent)
#save(birth_changes_table_long, file = "tables/birth_changes_table_long.Rdata")

birth_changes_table$N_total
```

## summary of facts for infobox
```{r, echo = F}
birth_change_fact <- birth_changes_table_long %>%
  filter(Birth_Changes == "Any change to birth plan" & Delivery_date == "After March 15, 2020") %>%
  mutate(Percent= round(Percent)) 

birth_change_fact$Percent

support_partner_fact <- birth_changes_table_long %>% 
  filter(Birth_Changes == "Support person not permitted at delivery" & 
           Delivery_date == "After March 15, 2020") %>%
  mutate(Percent = round(Percent))

support_partner_fact$Percent
```


## Types of changes to birth plan due to Covid-19
```{r, echo = F}
(birth_changes_table_long %>%
   filter(Birth_Changes != "Any change to birth plan" &
            Birth_Changes != "No changes" & Delivery_date == "After March 15, 2020")) %>%
  select(Birth_Changes, N_endorsed, Percent)
```

```{r,  echo = F}
birth_plot <- birth_changes_table_long %>%
      filter(Delivery_date == "After March 15, 2020" & 
               Birth_Changes != "No changes") %>%
      mutate(Birth_Changes = reorder(Birth_Changes, Percent)) %>%
      ggplot(aes(x = Birth_Changes, y = Percent)) +
      geom_bar(stat =  "identity", position = position_dodge(), fill = "navy") + 
      coord_flip()  +  ylim(0, 100)+
           ylab("Percent") +xlab ("") + custom_theme +
      labs( #title =Changes to Birth & Delivery during the Covid-19 outbreak", 
            caption = paste("Data aggregated from", birth_changes_table$N_total[birth_changes_table$Delivery_date =="After March 15, 2020"], "women who delivered after March 15th, 2020"))
        
    birth_plot
```

## Changes over time, based on delivery date
```{r}
week_interval <- function(date) {
  floor_date(date, "week")
}
month_interval <- function(date) {
  floor_date(date, "month")
}


birth_time_df <- mom_data_all %>%
  mutate(child_birth_date = ymd(child_birth_date),
         child_birth_month = month_interval(child_birth_date), 
         child_birth_week = week_interval(child_birth_date)) %>%
    # only include people who answered this Q 
  filter(covid19_survey_for_new_mothers_complete ==2) %>%
   # exclude people who did not answer questions
  filter(!is.na(any_birth_change) & 
           !is.na(child_birth_month)) 

birth_time_table <- birth_time_df %>%
  group_by(child_birth_week) %>%
    filter(child_birth_week >= ymd("2020-01-01")) %>%
  summarize(N_total = n(), 
            'Any change to birth plan' = sum(any_birth_change, na.rm = T),
            'Reduced access to preferred medications' = sum(birth_plan_changes___1 == 1, na.rm = T),
            'Changes to birth location' = sum(birth_plan_changes___2 == 1, na.rm = T),
            'Changes to birth schedule' = sum(birth_plan_changes___3 == 1 | 
                                                birth_plan_changes___4 == 1, na.rm = T),
            "Health care provider was not available" = sum(birth_plan_changes___5 == 1, na.rm = T),
            "Support person not permitted at delivery"= sum(birth_plan_changes___6 == 1, na.rm = T),
            'Separated from baby after delivery' = sum(birth_plan_changes___7 == 1 |
                                                         birth_plan_changes___8 == 1, na.rm = T),
            'Other change to birth plan' = sum(birth_plan_changes___10 == 1, na.rm = T),
            'No changes' = sum(birth_plan_changes___9 == 1, na.rm = T))

birth_time_table_long <- birth_time_table %>%
  gather(key = "Birth_Changes", value = "N_endorsed", -child_birth_week, -N_total) %>%
  mutate(Percent = round((N_endorsed/N_total)*100, 2)) %>%
  select(Birth_Changes, child_birth_week, N_endorsed, N_total, Percent) %>%
  filter(Birth_Changes != "No changes" & Birth_Changes != "Other change to birth plan" 
         & Birth_Changes != "Reduced access to preferred medications")


```


```{r}
## stuff needed
birth_change_time_N <- sum(birth_time_table_long %>%
  filter(Birth_Changes == "Any change to birth plan") %>% 
  select(N_total))
list_birth_changes <- unique(birth_time_table_long$Birth_Changes)

# plot
birth_time_table_long %>%
  # only include months with more than 10 respondents.
  filter(N_total > 10) %>%
#  filter(Birth_Changes == input$change_choice) %>%
  ggplot(aes (x = child_birth_week, y = Percent)) + 
  geom_line() + custom_theme + xlab("") + facet_wrap(~Birth_Changes, 
                                                     ncol = 2) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(caption = paste("Data aggregated from", birth_change_time_N, "women, with delivery dates binned in 2-week intevals."))

```