---
title: 'COPE data for NYU Langone cohort: Postnatal changes'
author: "Michelle.VanTieghem"
date: "June 2, 2020"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  word_document:
    toc: yes
---


# README 
Includes all participants who completed New Mom questionnaire at either baseline or Week 2 follow-up. \ Questions were optional so not all Qs have same  # of participants. \
Data collected between 04-04-2020 and 06-02-2020. \

```{r, echo = F, message = F, warning = F, include = F}
knitr::opts_chunk$set(message = F, warning = F)

# load packages
library(tidyverse)
library(ggplot2)
library(readxl)

# load custom packages
source(paste0(data_dir, "custom_functions.R"))

mom_data_all <- read.csv(paste0(data_dir, "clean_data/cleaned_new_mom_survey_2020-07-14.csv"))
(N_completed_maternal_Q <- nrow(mom_data_all))

```


  
# Changes in Postnatal health care due to Covid-19 outbreak
(check all that apply)
```{r, echo = F, include = F}
# get N for how many people answered this Q, to calculate %
N_total <- sum(!is.na(mom_data_all$any_postnatal_change))

# make table of changes
postnatal_changes_table <- mom_data_all %>%
  filter(covid19_survey_for_new_mothers_complete == 2) %>%
  # only include people who answered this Q
  filter(!is.na(any_postnatal_change)) %>%
  summarize(N_total = n(), 
            'Any change to postnatal care plan' = sum(any_postnatal_change, na.rm = T),
             'Family/friends unable to visit' = sum(postnatal_plan_changes___1 == 1, na.rm = T),
            'No access to lactation support' = sum(postnatal_plan_changes___2 == 1, na.rm = T),
            'Cancelled postpartum visit' = sum(postnatal_plan_changes___3 == 1, na.rm = T),
            'Virtual postpartum visit' = sum(postnatal_plan_changes___4 == 1, na.rm = T),
            "Unable to access preferred contraception" = sum(postnatal_plan_changes___5 == 1, na.rm = T),
            "Unable to talk about baby blues"= sum(postnatal_plan_changes___6 == 1, na.rm = T),
            'Virtual Baby well visits' = sum(postnatal_plan_changes___7 == 1, na.rm = T),
            'Cancelled Baby well visits' = sum(postnatal_plan_changes___8 == 1, na.rm = T),
            'Postponed immunizations for baby' = sum(postnatal_plan_changes___9 == 1, na.rm = T), 
            "Other change" =sum(postnatal_plan_changes___11 == 1, na.rm = T),
            "No change" = sum(postnatal_plan_changes___10 == 1, na.rm = T))

postnatal_changes_table_long <- postnatal_changes_table %>%
  gather(key = "Postnatal_Changes", value = "N_endorsed", -N_total) %>%
  select(Postnatal_Changes,N_endorsed, N_total) %>%
  mutate(Percent = round((N_endorsed/N_total)*100, 2)) %>%
  select(Postnatal_Changes,  N_endorsed, Percent)
```



## Types of changes to birth plan due to Covid-19
```{r, echo = F}
postnatal_changes_table_long %>%
   filter(Postnatal_Changes != "Any change to postnatal care plan" &
            Postnatal_Changes != "No change" ) %>%
  select(Postnatal_Changes, N_endorsed, Percent)

#save(postnatal_changes_table_long, file = "tables/postnatal_changes_table_long.Rdata")
```

```{r}
# create the plot
 postnatal_plot <- postnatal_changes_table_long %>%
      filter(Postnatal_Changes != "No change") %>%
      mutate(Postnatal_Changes = reorder(Postnatal_Changes, Percent)) %>%
      ggplot(aes(x = Postnatal_Changes, y = Percent)) +
      geom_bar(stat =  "identity", position = position_dodge(), fill = "navy") + 
      coord_flip()  + theme_bw() + ylab("Percent") + 
       ylim(0, 100) + xlab("") + custom_theme +
      labs( #title = "Changes to Postnatal care during the Covid-19 outbreak", 
            caption = paste("Data aggregated from", postnatal_changes_table$N_total, "women with infants under 6 months of age"))
postnatal_plot

#ggsave(postnatal_plot, file = "../plots/postnatal_changes.png")
```


## Current support 
How well have you been supported by your pre- and postnatal care provider(s) ?
```{r, echo = F, include = F}
support_table <- mom_data_all %>%
  filter(covid19_survey_for_new_mothers_complete == 2) %>%
  filter(!is.na(mom_care_support_change)) %>%
  summarize(N = n(), 
           "Very well supported" = sum(mom_care_support == 1, na.rm = T),
           "Somewhat well supported" = sum(mom_care_support == 2, na.rm = T), 
           "Not very well supported"= sum(mom_care_support == 3, na.rm = T))

support_table_long <- support_table %>%
  gather(key = "Support_level", value = "N_endorsed",  -N) %>%
  mutate(Percent = round((N_endorsed/N) *100, 2)) %>%
  select(Support_level, N_endorsed, Percent)
```

```{r, echo = F}
(support_table_long)

```


## Change in perinatal care due to Covid-19 outbreak
Did the support you received from your pre- and postnatal care provider(s) change due to the Covid-19 outbreak?
```{r, echo = F, include = F}

support_table <- mom_data_all %>%
  filter(!is.na(mom_care_support_change)) %>%
  summarize(N = n(), 
           "Significantly worsened" = sum(mom_care_support_change == 1, na.rm = T),
           "Somewhat worsened" = sum(mom_care_support_change == 2, na.rm = T), 
           "No change"= sum(mom_care_support_change == 3, na.rm = T), 
           "Not relevant" = sum(mom_care_support_change == 4, na.rm = T), 
           "Somewhat improved" = sum(mom_care_support_change == 5, na.rm = T), 
           "Significantly improved" = sum(mom_care_support_change == 6, na.rm = T))

support_table_long <- support_table %>%
  gather(key = "change", value = "N_endorsed",  -N) %>%
  mutate(Percent = round((N_endorsed/N) *100, 2)) %>%
  select(change, N_endorsed, Percent)
```



```{r, echo = F}

support_provider_plot <- support_table_long %>%
  mutate(change = reorder(change, Percent)) %>%
  filter(change != "Not relevant") %>%
ggplot(aes(x = change, y = Percent)) + 
  geom_bar(stat =  "identity", fill = "navy") + 
  ylim(0, 50) + coord_flip()  + xlab("") +
    ylab ("Percent") + custom_theme +
  labs(#title = "Changes in support from perinatal care providers during the Covid-19 outbreak",
       caption = paste("Data aggregated from", support_table$N, "women with infants under 6 months of age"))

support_provider_plot

#ggsave(support_provider_plot, file = "../plots/postnatal_provider_changes_in_support.png")
```

## ADD CONCERNS

## summary facts 
```{r, echo = F}
worsened_care_change <- support_table_long %>%
  filter(change == "Significantly worsened" | change == "Somewhat worsened") %>%
  summarize(Percent = round(sum(Percent)))

worsened_care_change$Percent


postnatal_change <- postnatal_changes_table_long %>%
   filter(Postnatal_Changes == "Any change to postnatal care plan") %>%
  mutate(Percent = round(Percent)) %>%
  select(Percent)

postnatal_change$Percent
```


## how did mom's concerns change over time
need to make a variable based on month they filled out this Q.
```{r, eval = F, include = f}

        #   mom mat care concern
         #  child health concern

#mom_concerns_table_long <- mom_concerns_table %>%
 # gather(key= "Concern", value = "mean_rating", -N_total, -child_birth_week)

mom_concerns_table_long %>%
  ggplot(aes (x = child_birth_week, y = mean_rating, 
              color= Concern, fill = Concern )) + 
  #geom_jitter(alpha = 0.5, height = 0.2, width = 0.2) +
  geom_smooth() + custom_theme + xlab("") +
  ylab("Mean Distress Rating") + ylim(0, 6) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(caption = paste("Data aggregated from", 
                       birth_change_time_N, 
                       "women, with delivery dates binned in 2-week intevals."))

```