---
title: "COPE EDA: impact & disruption"
author: "Michelle.VanTieghem"
date: "June 24, 2020"
output:
  html_document:
    number_sections: no
    df_print: kable
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---


```{r, echo = F, message = F, warning = F, include = F}
knitr::opts_chunk$set(message = F, warning = F)

library(psych)
# load packages
library(corrplot)
library(tidyverse)
library(ggplot2)
library(readxl)


# load custom packages
source(paste0(data_dir, "custom_functions.R"))
#data_dir <- "../"
baseline_df <- read.csv(paste0(data_dir, "clean_data/cleaned_baseline_survey_2020-07-14.csv"))

```

get relevant demo data 
```{r}
disrupt_df <- baseline_df %>%
   # making categorical
  mutate(currently_pregnant = as.factor(currently_pregnant), 
         treatment_any_past___1 = as.factor(treatment_any_past___1),
         treatment_mh_current___1 = as.factor(treatment_mh_current___1))
```

# material concern variables
0	Not of concern \
1	Mildly distressing \
2	Moderately distressing \
3	Highly distressing \

```{r}

concern_long <- disrupt_df %>%
  select(contains("concern"), status, first_child) %>%
  rename("Food or goods" = concern_food, 
         "Medicine" = concern_medicine, 
         "Baby supplies" = concern_babysupplies, 
         "Mental health care" = concern_mh, 
         "Health care" = concern_health_care,
         "Social interactions" = concern_social) %>%
  gather(key = "Concern", value = "Rating", -status, -first_child )

concern_long_table <- concern_long %>%
  filter(!is.na(Concern)) %>%
  group_by(status, first_child, Concern) %>%
  summarize(N_total = n(), 
            mean_rating = mean(Rating, na.rm = T), 
            upper_rating = mean_rating + sd(Rating, na.rm = T),
            lower_rating = mean_rating - sd(Rating, na.rm = T))
write.csv(concern_long_table, file = paste0(data_dir, "tables/concern_ratings_by_4groups.csv", row.names = F))
```

plot 
```{r}
N_total <- concern_long_table$N_total[concern_long_table$status == "New Mom"] +
  concern_long_table$N_total[concern_long_table$status == "Pregnant"]

list_concerns <- unique(concern_long_table$Concern)

# plot ratings from lowest to highest
concern_plot <- concern_long_table %>%
  mutate(Concern = fct_reorder(Concern, mean_rating)) %>%
#  filter(Concern %in% choice_concerns) %>%
  ggplot(aes(x = Concern, y = mean_rating, fill = first_child)) +  
    geom_bar(stat =  "identity", position = position_dodge()) + 
    theme_bw() + xlab("") + ylab("Concern Rating (mean)") +
      custom_theme + my_colors_fill + ylim(0, 3) + 
  coord_flip() + facet_grid(~status) +
  labs("", subtitle = "",
        fill = "First Child",
         caption ="0 = no concern, 3 = very concerned \n",
         paste("Data aggregated from", N_total, "new and expecting mothers"))

#Concerned about future access to...
concern_plot
##ggsave(concern_plot, file= paste0(data_dir, "plots/concern_ratings_by_preg_status.png")

```

# Covid Distress variables

```{r}
covid_distress <- disrupt_df %>%
  select( covid_family_distress, covid_self_distress, status, first_child) %>%
  drop_na() %>%
  rename("Covid illness: family" = covid_family_distress, 
         "Covid illness: self" = covid_self_distress)

covid_distress_long <- covid_distress %>%
  gather(key = "Distress", value = "Rating", -status, -first_child) %>%
  mutate(Distress = str_remove(Distress, "_distress")) %>%
  filter(!is.na(Distress)) %>%
  group_by(status, first_child, Distress) %>%
  summarize(N_total = n(), 
            mean_rating = mean(Rating, na.rm = T), 
            upper_rating = mean_rating + sd(Rating, na.rm = T),
            lower_rating = mean_rating - sd(Rating, na.rm = T))

write.csv(covid_distress_long, file = paste0(data_dir, "tables/covid_distress_by_4groups.csv", row.names = F))
```

plot ratings from lowest to highest
```{r}

N_total <- nrow(covid_distress)
covid_distress_plot <- covid_distress_long %>%
  ggplot(aes(x = Distress, y = mean_rating, fill = first_child)) + 
    geom_bar(stat =  "identity", position = position_dodge()) + 
    theme_bw() + xlab("") + ylab("Mean Distress Rating") +
      custom_theme + my_colors_fill + ylim(0, 6) + 
  facet_grid(~status) +
 labs(title = "Distress about potential covid-19 illness", 
       fill = "First Child",
         subtitle = "0 = no distress, 6 = extremely distressed",
       caption = paste("Data aggregated from", N_total, "new and expecting mothers"))

covid_distress_plot
#ggsave(covid_distress_plot, file= paste0(data_dir, "plots/covid_distress_ratings_by_preg_status.png", width = 9, height = 5))
```


# Financial Distress variables

```{r}
job_distress <- disrupt_df %>%
  select(current_job_distress, future_job_distress, status, first_child) %>%
  drop_na() %>%
  rename("Current" = current_job_distress, 
         "Future" = future_job_distress)

job_distress_long <- job_distress %>%
  gather(key = "Distress", value = "Rating", -status, -first_child) %>%
  mutate(Distress = str_remove(Distress, "_distress")) %>%
  filter(!is.na(Distress)) %>%
  group_by(status, first_child, Distress) %>%
  summarize(N_total = n(), 
            mean_rating = mean(Rating, na.rm = T), 
            upper_rating = mean_rating + sd(Rating, na.rm = T),
            lower_rating = mean_rating - sd(Rating, na.rm = T))

write.csv(job_distress_long, file = paste0(data_dir, "tables/job_distress_by_4groups.csv", row.names = F))
```

plot ratings from lowest to highest
```{r}

N_total <- nrow(job_distress)
job_distress_plot <- job_distress_long %>%
  ggplot(aes(x = Distress, y = mean_rating, fill = status)) + 
    geom_bar(stat =  "identity", position = position_dodge()) + 
    theme_bw() + xlab("Job Worries") +
  ylab("Mean Distress Rating") +
      custom_theme + my_colors_fill + ylim(0, 6) + 
 labs(title = "", 
         subtitle = "",
       caption ="(0 = no distress, 6 = extremely distressed)")

job_distress_plot
#ggsave(job_distress_plot, file= paste0(data_dir, "plots/job_distress_ratings_by_preg_status.png")
```

# Disrupt variables
```{r}
disrupt_long <- disrupt_df %>%
  select(contains("disrupted"), status, first_child, -disrupted_meal_rating) %>%
   drop_na() %>%
  rename("work activities" = disrupted_work_rating, 
         "social activities" = disrupted_social_activities_rating, 
         "physical activities" = disrupted_physical_rating)


disrupt_long_table <- disrupt_long %>%
  gather(key = "disrupt", value = "Rating", -status, -first_child) %>%
  group_by(status, first_child, disrupt) %>%
  summarize(N_total = n(), 
            mean_rating = mean(Rating, na.rm = T), 
            upper_rating = mean_rating + sd(Rating, na.rm = T),
            lower_rating = mean_rating - sd(Rating, na.rm = T))

write.csv(disrupt_long_table, file = paste0(data_dir, "tables/disruption_ratings_by_4groups.csv", row.names = F))

```

plot 
```{r}
N_total <- nrow(disrupt_long)
# plot ratings from lowest to highest
disrupt_plot <- disrupt_long_table %>%
  mutate(disrupt = fct_reorder(disrupt, mean_rating)) %>%
  ggplot(aes(x = disrupt, y = mean_rating, fill = first_child)) +  
    geom_bar(stat =  "identity", position = position_dodge()) + 
    theme_bw() + xlab("") + ylab("Mean Disruption Rating") + 
      custom_theme + my_colors_fill + ylim(0, 6) + coord_flip() +
  labs(title = "Disruptions due to the Covid-19 pandemic",
        fill = "First Child",
         caption = " (0 = none, 6 = extreme disruption)") + 
    facet_grid(~status)
disrupt_plot
##ggsave(disrupt_plot, file= paste0(data_dir, "plots/disrupt_ratings_by_preg_status.png", width = 9, height = 5))
```


## covid valence impact

```{r}

valence_table <- disrupt_df %>% 
  select(record_id, first_child, impact_covid_valence, status) %>%
  mutate(valence = ifelse(impact_covid_valence> 0, "positive",
                          ifelse(impact_covid_valence == 0, "neutral", "negative"))) %>%
    filter(!is.na(valence)) %>%
  group_by(status, first_child) %>%
  summarize(N_total = n(),
            negative = sum(valence == "negative", na.rm = T),
            positive = sum(valence == "positive", na.rm = T),
            neutral = sum(valence == "neutral", na.rm = T))

valence_long <- valence_table %>%
  gather(key = "valence", value = "N_endorsed", -N_total, -status, -first_child) %>%
  mutate(percent = round((N_endorsed/N_total)*100))

write.csv(valence_long, file = paste0(data_dir, "tables/valence_ratings_by_4groups.csv", row.names = F))
```

plot
```{r}
valence_plot <- valence_long %>%
   mutate(valence = fct_reorder(valence, percent)) %>%
  ggplot(aes(x = valence, y = percent, fill = first_child)) + coord_flip()  +
  geom_bar(stat = "identity", position = position_dodge()) + 
   xlab ("") + facet_grid(~status) +
  ylab("Percent") +custom_theme + my_colors_fill + my_colors_color +
  labs(title = "", 
        fill = "First Child",
          caption = paste("Data aggregated from", N_total, "new and expecting mothers"))

valence_plot
#ggsave(valence_plot, file= paste0(data_dir, "plots/valence_impact_on_life_by_preg_status.png")
```



## General impact of covid
```{r}
impact <- baseline_df %>%
  select(status, first_child, impact_covid_rating, stress_covid_rating) %>%
  # put on scale so 0 = no stress or no impact.
  mutate(impact_covid_rating = impact_covid_rating -1, 
         stress_covid_rating = stress_covid_rating -1)

impact_long <- impact %>%
  gather(key = "measure", value = "Rating", 
         impact_covid_rating, stress_covid_rating,  -first_child, -status) %>%
  mutate(measure = ifelse(measure == "impact_covid_rating", 
                          "Daily Impact", "Stress Level"))

impact_long_table <- impact_long %>%
  group_by(status, first_child, measure) %>%
  summarize(mean_rating = mean(Rating, na.rm = T), 
            upper_rating = mean_rating + sd(Rating, na.rm = T),
            lower_rating = mean_rating - sd(Rating, na.rm = T))

write.csv(impact_long_table, file = paste0(data_dir, "tables/impact_ratings_by_4groups.csv", row.names = F))
```

plot
```{r}
# plot ratings from lowest to highest
impact_plot <- impact_long_table %>%
  ggplot(aes(x = measure, y = mean_rating, fill = first_child)) +  
    geom_bar(stat =  "identity", position = position_dodge()) + 
    xlab("") + ylab("Mean Rating") +
   custom_theme + my_colors_fill + ylim(0, 6) +
     labs(title = "Self-reported Rating", subtitle = "0 = nothing, 6 = extreme",
           fill = "First Child") +
   facet_grid(~status)

impact_plot
##ggsave(impact_plot, file= paste0(data_dir, "plots/impact_stress_ratings_by_preg_status.png")
```

## biggest stressor
```{r}
greatest_stress <- baseline_df %>%
  select(status,first_child, source_stress_covid) %>%
  filter(!is.na(source_stress_covid)) 

stress_table <- greatest_stress %>%
  group_by(status, first_child) %>%
  summarize(N_total = n(), 
            "Health concerns" = sum(source_stress_covid == 1), 
            "Financial concerns" = sum(source_stress_covid == 2), 
            "Impact on your child" = sum(source_stress_covid == 3), 
            "Impact on your partner" = sum(source_stress_covid == 4),
            "Impact on your community" = sum(source_stress_covid == 5),
            "Impact on family members"= sum(source_stress_covid == 6), 
            "Impact on close friends" = sum(source_stress_covid ==7),
            "Impact on society" = sum(source_stress_covid == 8),
            "Access to food" = sum(source_stress_covid == 9),
            "Access to baby supplies" = sum(source_stress_covid == 10),
            "access to mental health care" = sum(source_stress_covid == 11), 
            "Social distancing" = sum(source_stress_covid == 12),
            "Other" = sum(source_stress_covid == 13))
            

stress_long <- stress_table %>%
  gather(key = "source_stress", value = "N_endorsed", -status, -first_child, -N_total) %>% 
  mutate(percent = round((N_endorsed/N_total)*100)) %>%
  filter(percent > 7)

write.csv(stress_long, file = paste0(data_dir, "tables/greatest_source_stress_by_4groups.csv", row.names = F))
```

plot
```{r}

# plot ratings from lowest to highest
stress_plot <- stress_long %>%
    mutate(source_stress = fct_reorder(source_stress, percent)) %>%
  ggplot(aes(x = source_stress, y = percent, fill = first_child)) +  
    geom_bar(stat =  "identity", position = position_dodge()) + 
    theme_bw() + ylab("Percent") + xlab("") + coord_flip() + 
      custom_theme + my_colors_fill + facet_grid(~status) +
  labs(#title = "Single greatest source of stress due to Covid-19", 
         fill = "First Child",
         caption = paste("Data aggregated from", N_total, "new and expecting mothers"))
stress_plot

#ggsave(stress_plot, file= paste0(data_dir, "plots/single_greatest_stress_by_preg_status.png")
```
