---
title: "COPE data cleaning for New Mom survey"
author: "Michelle.VanTieghem"
date: "June 2, 2020"
output:
  html_document:
    number_sections: no
    df_print: kable
    code_folding: show
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      
---



```{r, message = F, warning = F, include = F, echo = F}
knitr::opts_chunk$set(message = F, warning = F)

library(tidyverse)
library(ggplot2)
library(readxl)
```

# README
cleaning new mother COPE survey data for further analysis \
output: clean_data/new_mom_df_clean.Rda" \
combines across baseline and follow-up, depending on when gave birth \
removes duplicates and reformats, only keeps completed surveys

# load data
## load cleaned baseline data from script 1a
```{r}
load("../cope_data_dashboard/clean_data/baseline_deidentified_df.Rdata")

```

## load Week 2 follow-up data (longitudinal project)
```{r, echo = F, message = F}
# most recent version: downloaded 07-02-2020
long_df <- read_csv("../raw_data/survey_data/COPELongitudinalStud_DATA_use.csv") %>%
  select(-contains("timestamp"))  %>%
   mutate(child_birth_date = as.character(child_birth_date), 
         birth_date = as.character(birth_date), 
         today_date = as.character(today_date),
         pregnant_due_date = as.character(pregnant_due_date),
         covid_tested_neg_date = as.character(covid_tested_neg_date),
         date_of_visit = as.character(date_of_visit),
        status = ifelse(is.na(currently_mother) & is.na(birth_update), "Pregnant", "New Mom"))%>%
  rename(baseline_date = today_date,
         fu_date = date_of_visit) %>%
  mutate(phase = "phase2_long")

# first follow-up - retain people who reported giving birth at followup
fu1_df <-long_df %>%
  filter(redcap_event_name == "covid_week_2_follo_arm_1" & birth_update == 1) %>%
  select(-child_birth_date) %>%
  # get the follow-up birth date and rename it.
  rename(child_birth_date = child_birth_date2)
nrow(fu1_df)

# second follow-up - keep people who reported giving birth
fu2_df <-long_df %>%
  # then filter only for this event, for newly delivered, complete surveys.
  filter(redcap_event_name == "covid_week_4_follo_arm_1" &
           birth_update == 1 & covid19_survey_for_new_mothers_complete == 2) %>%
  select(-child_birth_date) %>%  # all NAs...
  rename(child_birth_date = child_birth_date2) 
nrow(fu2_df)

# get birth dates from the fu1 deliveries that were missing from fu2 
fu1_df <- fu2_df %>% 
  filter(fu2_df$record_id %in% fu1_df$record_id) %>%
  select(record_id, child_birth_date) %>%
  left_join(fu1_df, ., by = c("record_id")) %>%
  # if the birth date was missing in fu1, get it from fu2
  mutate(child_birth_date = ifelse(is.na(child_birth_date.x), 
                                   as.character(child_birth_date.y), 
                                   as.character(child_birth_date.x))) %>%
  select(-child_birth_date.x, -child_birth_date.y)
nrow(fu1_df)
```

## clean new mom data (pooling across baseline + follow ups)
```{r, echo = F}
# get demo data for all moms 
demo_data <- baseline_clean2 %>%
  select(record_id, baseline_date, 
         currently_pregnant, currently_mother,
        income_brackets, median_income, education,
         starts_with("race"), ethnicity,
         # covid testing
        contains("test"), contains ("symp"),
         #pull items of interest
        contains("treat"),contains("concern"), contains("distress"),
         contains("social"), contains("drug"), contains("opiod"),
         starts_with("bsi"),  starts_with("ptsd"), 
        # make surre mom/ preg concerrn ratingns don't get pulledin
        -contains("mom"), -contains("preg"))

# extract only the mom survey part for the moms who filled out at baseline
mom_data <- baseline_clean2 %>%
 filter(status == "New Mom" & covid19_survey_for_new_mothers_complete == 2) %>%
  select(record_id, baseline_date,  child_birth_date, 
         status,  starts_with("birth_plan"), starts_with("mom"), 
         starts_with("postnatal"), covid19_survey_for_new_mothers_complete) %>%
  mutate(birth_update = NA, mom_medical_vitamins = NA) %>%
  rename(new_mom_survey_date = baseline_date)
nrow(mom_data)

# who gave birth between baseline and FU1?
mom_data1 <- fu1_df %>%
  filter(birth_update == 1 & covid19_survey_for_new_mothers_complete == 2) %>%
    filter(!record_id %in% mom_data$record_id) %>%
  mutate(status = "New Mom") %>%
  select(record_id, fu_date, child_birth_date, status, birth_update,
         starts_with("birth_plan"), starts_with("mom"), 
         starts_with("postnatal"), covid19_survey_for_new_mothers_complete) %>%
  rename(new_mom_survey_date = fu_date)

# who gave birth between FU1 and FU2? 
mom_data2 <- fu2_df %>%
  filter(birth_update == 1 & covid19_survey_for_new_mothers_complete == 2) %>%
  # remove people who already did new mom Q at baseline or FU1, 
  # and accidentally did the Q a second time.
  filter(!record_id %in% mom_data$record_id & 
           !record_id %in% mom_data1$record_id) %>%
  mutate(status = "New Mom") %>%
  select(record_id, fu_date, child_birth_date, status, birth_update,
         starts_with("birth_plan"), starts_with("mom"), 
         starts_with("postnatal"), covid19_survey_for_new_mothers_complete) %>%
  rename(new_mom_survey_date = fu_date)

```

## combine mom data
```{r, echo = F}

merge_names <- names(mom_data)[! names(mom_data) %in% names(mom_data2)]

mom_data_all <- rbind(mom_data, mom_data1, mom_data2) %>%
  merge(., demo_data, by = "record_id") 
length(unique(mom_data_all$record_id))

nrow(mom_data) + nrow(mom_data1) + nrow(mom_data2)

# check that no duplicate participants after merging
identical(length(unique(mom_data_all$record_id)), nrow(mom_data_all))
```

## Add coding for variables of interest
```{r, echo = F, include = F}

mom_data_all <- mom_data_all %>%
  mutate(During_covid = as.factor(ifelse(child_birth_date >= as.Date("2020-03-15") |
                     # some moms didn't report child birth in the week 2 update
                                           !is.na(birth_update), 1, 0)),
      Delivery_date = ifelse(During_covid == 0, 
         "Before March 15, 2020", "After March 15, 2020"),
         # because data is spread across 10 dif answers
     # make indicator for whether any change was endorsed 
       no_resp_birth_change = ifelse((birth_plan_changes___1 == 0) &
                               (birth_plan_changes___2 == 0) &
                                 (birth_plan_changes___3 == 0) &
                                 (birth_plan_changes___4 == 0) &
                                 (birth_plan_changes___5== 0) &
                                 (birth_plan_changes___6== 0) &
                                 (birth_plan_changes___7== 0) &
                                 (birth_plan_changes___8== 0) &
                                 (birth_plan_changes___9== 0) &
                                 (birth_plan_changes___10 == 0), 1, 0),
     any_birth_change = ifelse(birth_plan_changes___1 == 1 |
                               birth_plan_changes___2 == 1 |
                               birth_plan_changes___3 == 1 |
                               birth_plan_changes___4 == 1 |
                               birth_plan_changes___5 == 1 |
                               birth_plan_changes___6 == 1 |
                               birth_plan_changes___7 == 1 |
                               birth_plan_changes___8 == 1 |
                                 # option 9 = no change
                               birth_plan_changes___10 == 1, 1, 
                                      # mark people who answered none of these options as NA.
                               ifelse(no_resp_birth_change == 1, NA, 0)),
     # make prettier version for plotting  
     Birth_change_factor = as.factor(ifelse(any_birth_change == 0, "No changes", "Changes to Birth Plan")),

    no_resp_postnatal_change = ifelse(postnatal_plan_changes___1 == 0 &
                                     postnatal_plan_changes___2 ==0 &
                                   postnatal_plan_changes___3 == 0 &
                                  postnatal_plan_changes___4 == 0 &
                                  postnatal_plan_changes___5 == 0 &
                                   postnatal_plan_changes___6 == 0 &
                                   postnatal_plan_changes___7 == 0 &
                                    postnatal_plan_changes___8 == 0 &
                                     postnatal_plan_changes___9 == 0 &
                                    postnatal_plan_changes___10 == 0 &
                                     postnatal_plan_changes___11 == 0, 1, 0),
     any_postnatal_change = ifelse(postnatal_plan_changes___1 == 1 |
                                     postnatal_plan_changes___2 == 1 |
                                   postnatal_plan_changes___3 == 1 |
                                  postnatal_plan_changes___4 == 1 |
                                  postnatal_plan_changes___5 == 1 |
                                   postnatal_plan_changes___6 == 1|
                                   postnatal_plan_changes___7 == 1 |
                                    postnatal_plan_changes___8 == 1 |
                                     postnatal_plan_changes___9 == 1 |
                                    # option 10 = no change
                                     postnatal_plan_changes___11 == 1, 1,
                                  # mark people who answered none of these options as NA.
                                   ifelse(no_resp_birth_change == 1, NA, 0)),
     # make prettier version for plotting  
     Postnatal_change_factor = as.factor(ifelse(any_postnatal_change == 0, "No changes", "Changes to Postnatal care")))
```

### check variable coding
```{r}

summary(as.factor(mom_data_all$no_resp_birth_change))
summary(as.factor(mom_data_all$any_birth_change))

summary(as.factor(mom_data_all$no_resp_postnatal_change))
summary(as.factor(mom_data_all$any_postnatal_change))
```


## save cleaned data 
```{r}

save(mom_data_all, file ="../cope_data_dashboard/clean_data/new_mom_df_clean.Rdata")
```